VOI-based Monte Carlo Sampling

### 2011-12-07 paper/vago/paper.tex
better bounds on blinkered voi
initial commit

@@

Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
\begin{description}
\item[Hoeffding's inequality \rm{\cite{Hoeffding.ineq}}:] 
\begin{equation}
\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
\label{eqn:conc-hoeffding}
\end{equation}
\item[Empirical Bernstein's inequality
  \rm{\cite{MaurerPontil.benrstein}}:]\footnote{see
    Appendix~\ref{app:deriv-conc-empbernstein} for derivation}
\begin{eqnarray}
\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
                    &\le& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
\label{eqn:conc-empbernstein}
\end{eqnarray}
where sample variance $\overline\sigma_n^2$ is
\begin{equation}
\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
\label{eqn:sample-variance}
\end{equation}
\end{description}
Bounds (\ref{eqn:conc-hoeffding}, \ref{eqn:conc-empbernstein}) are symmetrical
around the mean. Bound~(\ref{eqn:conc-empbernstein}) is tighter than
(\ref{eqn:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.

@@

\begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
  bounded from above as
\begin{equation}
  \Lambda_i^p \le \left\{
  \begin{array}{l l}
    \Pr(\IE[X_i] \le \overline X_\beta)\overline X_\beta & \mbox{if $i=\alpha$} \\
    \Pr(\IE[X_i] \ge \overline X_\alpha)(1-\overline X_\alpha) & \mbox{otherwise}
  \end{array} \right.
\label{eqn:thm-vopi}
\end{equation}
\end{thm}

@@

\begin{thm} Denote the current number of samples of the $i$th arm as
  $n_i$. The blinkered estimate $\Lambda_i^b$ of intrinsic value of
  information of pulling the $i$th arm for the remaining budget of $N$
  samples is bounded from above as
\begin{equation}
  \Lambda_i^b \le \left\{
  \begin{array}{l l}
    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i} & \mbox{if $i=\alpha$} \\
    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline  X_\alpha)\frac N {N+n_i} &  \mbox{otherwise}
  \end{array} \right.
\label{eqn:thm-be}
\end{equation}
where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ 
samples.
\end{thm}

 The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) can be bounded from above using concentration
inequalities. In particular, Lemma~\ref{lemma:hoeffding-prob-bounds} is
based on the Hoeffding inequality (\ref{eqn:conc-hoeffding}):
\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) are bounded from above as
\begin{eqnarray}
\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)& \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
\Pr(\overline X_i' \le \overline X_\beta|i=\alpha)& \le &  2\exp\left(-\frac {2(\overline X_i - \overline X_\beta)^2  (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2}  \right) < 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2} \right) <  2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
\label{eqn:probound-perf-hoeffding}
\end{eqnarray}
\label{lemma:hoeffding-prob-bounds}
\end{lmm}

@@

Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
\Lambda$ on the perfect value of information should be pulled at each
step. This way, arms known to have a low VOI  would be pulled less frequently.

@@

The upper bounds~(\ref{eq:lambda-hoeffding-bounds}, \ref{eq:lambda-bernstein-bounds}) decrease exponentially with the
number of pulls $n$. When the upper bound of the VOI for all arms
becomes lower than a threshold $\lambda$, which can be chosen based on
resource constraints, the sampling should be stopped, and an arm should
be chosen.

@@

\section{Empirical Evaluation}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bests-colorful.pdf}
\caption{Go: best winning rate comparison}
\label{fig:best-winning-rate}
\end{figure}

@@

@@

Theorem~4 in~\cite{MaurerPontil.benrstein} states that
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
Therefore
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
\delta.\]
$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
square equation
\[a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge a)$,
gives
\[\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
Other derivations, giving slightly different results, are possible.

@@

The bound can be supposedly be improved by selecting a midpoint
$0 < y < \overline X_\beta$ and computing the bound as the sum of two parts:
\begin{itemize}
\item $\overline X_\beta-y$ multiplied by the probability that
  $\mu_\alpha \le \overline X_\beta$;
\item $\overline X_\beta$ multiplied by the probability that $\mu_\alpha\le
  y$.
\end{itemize}.
\[ V_{\overline X=\overline X_\alpha} \le (\overline X_\beta-y)\exp\left(-2n(\overline X_\alpha-\overline X_\beta)^2\right)+\overline X_\beta\exp\left(-2n(\overline X_\alpha-y)^2\right) \]

\vspace{\baselineskip}

The minimum of $V^*$ is achieved when $\frac {dV^*} {dy}=0$, that
is, when $y$ is the root of the following equation:
\[ 4\overline X_\beta n(\overline X_\alpha-y)=\exp\left(-2n\left(\frac {\overline X_\alpha-\overline X_\beta}
    {\overline X_\alpha-y}\right)^2\right) \]
If a root in the interval  $0\le y\le\beta$ exists, then the
number of samples is bounded as
\[n\le\frac 1 {4\overline X_\beta(\overline X_\alpha-\overline X_\beta)}\]
by observing that the right-hand side 
is at most $1$ (a negative power), and the left-hand side is at least
$4\overline X_\beta n(\overline X_\alpha-\overline X_\beta)$.
So, the bound can supposedly be improved for smaller values of
$n$. The improvement is more significant when the current best and
second-best sample means are close.

The derivation for the other case
(sampling an item that can be better than the current best) is
obtained by substitution $1-\overline X, 1-\overline X_\alpha, 1-y$ instead of
$\overline X_\alpha, \overline X_\beta, y$:

\[ V_{\overline X\ne\overline X_\alpha} \le (y-\overline X_\alpha)\exp\left(-2n(\overline X_\alpha-\overline
  X)^2\right)+(1-\overline X_\alpha)\exp\left(-2n(y-\overline X)^2\right) \]

The anticipated influence of the improved
VOI estimate would be that the selected item will be a less discovered
one and further from the current best or second-best.

A closed-form solution for $y$ cannot be obtained, but given
$\overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
computed. It should be determined empirically whether the improved
estimate has justified influence on the performance of the algorithm.

>>>>

\section{Related Work}

\textit{TODO}

@@

Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
\begin{description}
\item[Hoeffding's inequality \rm{\cite{Hoeffding.ineq}}:] 
\begin{equation}
\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
\label{eqn:conc-hoeffding}
\end{equation}
\item[Empirical Bernstein's inequality
  \rm{\cite{MaurerPontil.benrstein}} \textrm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
\begin{eqnarray}
\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
                    &\le& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
\label{eqn:conc-empbernstein}
\end{eqnarray}
where sample variance $\overline\sigma_n^2$ is
\begin{equation}
\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
\label{eqn:sample-variance}
\end{equation}
\end{description}
Bounds (\ref{eqn:conc-hoeffding}, \ref{eqn:conc-empbernstein}) are symmetrical
around the mean. Bound~(\ref{eqn:conc-empbernstein}) is tighter than
(\ref{eqn:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.

@@

\begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
  bounded from above as
\begin{equation}
  \Lambda_i^p \le \left\{
  \begin{array}{l l}
    \Pr(\IE[X_i] \le \overline X_\beta)\overline X_\beta & \mbox{if $i=\alpha$} \\
    \Pr(\IE[X_i] \ge \overline X_\alpha)(1-\overline X_\alpha) & \mbox{otherwise}
  \end{array} \right.
\label{eqn:thm-vopi}
\end{equation}
\label{thm:vopi}
\end{thm}

\begin{proof} For the case $i\ne \alpha$, the probability that the perfect
  information about the $i$th arm changes the final choice is
  $\Pr(\IE[X_i] \ge \overline X_\alpha)$. $\IE[X_i] \le 1$ by definition,
  therefore the maximum increase in the expected reward is
  $(1-\overline X_\alpha)$. Thus the intrinsic value of perfect
  information is at most $\Pr(\IE[X_i] \ge \overline
  X_\alpha)(1-\overline X_\alpha)$.
  Proof for the case $i=\alpha$ is similar.
\end{proof}

@@

\begin{thm} Denote the current number of samples of the $i$th arm as
  $n_i$. The blinkered estimate $\Lambda_i^b$ of intrinsic value of
  information of pulling the $i$th arm for the remaining budget of $N$
  samples is bounded from above as
\begin{equation}
  \Lambda_i^b \le \left\{
  \begin{array}{l l}
    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i}
       \le N\Pr(\overline X_i'\le\overline X_\beta)\frac {\overline X_\beta} {n_i} & \mbox{if $i=\alpha$} \\
    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline  X_\alpha)\frac N {N+n_i}
       \le N\Pr(\overline X_i'\ge\overline X_\alpha)\frac {(1-\overline  X_\alpha)} {n_i} & \mbox{otherwise}
  \end{array} \right.
\label{eqn:thm-be}
\end{equation}
where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ 
samples.
\label{thm:be}
\end{thm}

\begin{proof}
The proof is similar to the proof of Theorem~\ref{thm:vopi}, with
$\overline X'_i$ substituted instead of $\IE[X_i]$, and, if
$i\ne\alpha$, the upper bound on $X_i'$ is $X_\alpha+(1-X_\alpha)
\frac N {N+n_i}$; if $i=\alpha$, the lower bound on $X_i'$ is
$X_\beta - X_\beta \frac N {N+n_i}$.
\end{proof}

The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be})
can be bounded from above using concentration inequalities. In
particular, Lemma~\ref{lemma:hoeffding-prob-bounds} (proved in 
Appendix~\ref{app:hoeffding-prob-bounds-proof}) is based on the
Hoeffding inequality (\ref{eqn:conc-hoeffding}):
\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi},
\ref{eqn:thm-be}) are bounded from above as
\begin{eqnarray}
\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)&
   \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)&
   \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)
\label{eqn:probound-perf-hoeffding}
\end{eqnarray}
\vspace{-1.5\baselineskip}
\begin{eqnarray}
\Pr(\overline X_i' \le \overline X_\beta|i=\alpha) &
   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_i - \overline X_\beta)\right)^2 n_i \right)
   \le 2\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i \right) \nonumber\\
\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha) &
   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha - \overline X_i)\right)^2 n_i \right)
    \le 2\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i \right)
\label{eqn:probound-blnk-hoeffding}
\end{eqnarray}
\label{lemma:hoeffding-prob-bounds}
\end{lmm}

An upper bound on the intrinsic value of perfect information is obtained by substituting
(\ref{eqn:probound-perf-hoeffding}) into (\ref{eqn:thm-vopi}):
\begin{equation}
  \Lambda_i^p \le \hat\Lambda_i^p=\left\{
  \begin{array}{l l}
    \exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right)\overline X_\beta & \mbox{if $i=\alpha$} \\
    \exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)(1-\overline  X_\alpha)  &  \mbox{otherwise}
  \end{array} \right.
\label{eqn:bound-perf-hoeffding}
\end{equation}
An upper bound on the blinkered VOI estimate with ordering independent of $N$ is obtained
by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
\begin{equation}
  \Lambda_i^b \le \hat\Lambda_i^b=\left\{
  \begin{array}{l l}
    C \exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)\frac {\overline X_\beta} {n_i}
      & \mbox{if $i=\alpha$} \\
    C \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)\frac {(1-\overline  X_\alpha)} {n_i}
      &  \mbox{otherwise}
  \end{array} \right.
\label{eqn:bound-blnk-hoeffding}
\end{equation}
where $C=2N$. The bound is generally tighter when $N$ is small relative to $n_i$,
and the estimate $\hat\Lambda_i^b$ can be viewed as a relaxation of the myopic VOI
estimate to longer sequences of actions, such
that the VOI estimate is positive for some of the actions.

@@

Following the principles of rational metareasoning, an arm with
the highest VOI should be pulled at each step. The upper bounds
established in Theorems~\ref{thm:vopi},~\ref{thm:be} can be used
as VOI estimates. Blinkered VOI estimates~(\ref{eqn:thm-be},
 \ref{eqn:bound-blnk-hoeffding}) can be viewed
as approximations of VOI estimates for myopic or
receding horison policies. As illustrated by the empirical evaluation
(Section~\ref{sec:empirical-evaluation}), estimates based on upper
bounds on the VOI result in rational sampling policies, and exhibit
performance comparable to the performance of some state-of-the-art
heuristic algorithms.

@@

The simplest termination condition for a sampling policy is the
budget---a fixed number of samples performed before a decision is
made. When a sample has a cost commensurable with the value of
information of a measurement, an upper bound on the intrinsic VOI can
be used to stop the sampling if the intrinsic VOI of any action
is less than the cost of sampling $C$:
\begin{equation}
\mbox{stop if } \max_i \Lambda_i \le C
\end{equation}
Blinkered VOI estimates~(\ref{eqn:thm-be},
\ref{eqn:bound-blnk-hoeffding}) include the remaining budget $N$ as a
factor, but given the cost of a single sample $c$, the cost of the
remaining samples accounted for in estimating the intrinsic VOI is
$C=cN$. $N$ can be dropped on both sides of the inequality,
and a viable stopping condition is
\begin{equation}
\begin{array}{c r r l l}
&\frac 1 N \Lambda_\alpha^b
\le\Pr(\overline X_\alpha'\le\overline X_\beta)\frac {\overline  X_\beta} {n_\alpha}&\le& c
&\nonumber\\
\mbox{and}&&&&\nonumber\\
&\frac 1 N \max_i\Lambda_i^b
\le \max_i\Pr(\overline X_i'\ge\overline X_\alpha)\frac {(1-\overline  X_\alpha)} {n_i}&\le& c
 & \forall i: i\ne\alpha
\end{array}
\label{eqn:stopping-blnk}
\end{equation}
The empirical evaluation (Section~\ref{sec:empirical-evaluation})
confirms viability of this stopping condition and illustrates the
influence of the sample cost $c$ on the performance of
the sampling policy.

@@

\section{Empirical Evaluation}
\label{sec:empirical-evaluation}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bests-bw.pdf}
\caption{Go: best winning rate comparison}
\label{fig:best-winning-rate}
\end{figure}

\section{Summary and Future Work}

\textit{TODO}

@@

\section{Proof of Lemma~\ref{lemma:hoeffding-prob-bounds}}
\label{app:hoeffding-prob-bounds-proof}

\begin{proof}[]
Equations (\ref{eqn:probound-perf-hoeffding}) is a direct
application of the Hoeffding inequality (\ref{eqn:conc-hoeffding}).

Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
observation that if $i\ne\alpha$, $\overline X_i'>\overline X_\alpha$
if and only if the mean $\overline X_i^+$ of $N$ samples from $n_i+1$
to $N$ is at least $\overline X_\alpha+(\overline X_\alpha-\overline
X_i)\frac {n_i} N$.

For any $\delta$, the probability that $\overline X_i'$ is \textit{not
greater} than
$X_\alpha$ is more than the probability that $\IE[X_i]\le\overline
X_i+\delta$ \textit{and} $\overline X_i'\ge \IE[X_i]+\overline X_\alpha - \overline
  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$;
therefore, the probability that $\overline X_i'$ is \textit{greater}
than than $X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
X_i+\delta$ \textit{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline
  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$.
Using the union bound,
\begin{equation}
\Pr(\overline X_i'\ge \overline X_\alpha)\le
  \Pr(\IE[X_i]-\overline X_i \ge \delta) +
  \Pr(\overline X_i^+ - \IE[X_i] \ge \overline X_\alpha - \overline
  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N)
\end{equation}
Bounding the probabilities on the right side using the Hoeffding
inequality, obtain:
\begin{equation}
\Pr(\overline X_i'\ge \overline X_\alpha)\le
  \exp(-2\delta^2n_i)+\exp\left(-2\left((\overline X_\alpha - \overline
  X_i)\left(1+\frac {n_i} N\right) - \delta\right)^2N\right)
\end{equation}
$\delta^2n = -2\left((\overline X_\alpha - \overline X_i)
(1+\frac {n_i} N) - \delta\right)^2N$ when
$\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha
- \overline X_i) \ge (\sqrt 2 - 1)(\overline X_\alpha-\overline X_i)$,
therefore
\begin{eqnarray}
\Pr(\overline X_i'\ge \overline X_\alpha) 
& \le & 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_\alpha - \overline X_i)\right)^2 n_i\right)\nonumber \\
& \le & 2\exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
 \le 2\exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)
\end{eqnarray}
Derivation for the case $i=\alpha$ is similar.
\end{proof}

@@

Theorem~4 in~\cite{MaurerPontil.benrstein} states that
\begin{equation}
\Pr\left(\IE[X]-\overline X_n
    \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta
\end{equation}
Therefore
\begin{equation}
\Pr\left(\IE[X]-\overline X_n \ge
              \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n }
                      + \frac {7 \ln 2/\delta} {3(n-1)}\right)
   \le \delta.
\end{equation}
$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$
 is a root of square equation $a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0$
which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge a)$,
gives
\begin{equation*}
\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)
\end{equation*}
Other derivations, giving slightly different results, are possible.

@@

Bounds (\ref{eqn:bound-perf-hoeffding}, \ref{eqn:bound-blnk-hoeffding}) have the form
\begin{equation}
  \Lambda_i \le \hat\Lambda_i=\left\{
  \begin{array}{l l}
    \exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right)(\overline X_\beta-B) & \mbox{if $i=\alpha$} \\
    \exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)(A-\overline  X_\alpha)  &  \mbox{otherwise}
  \end{array} \right.
\label{eqn:bound-generic-hoeffding}
\end{equation}
where $A$ is the upper bound on the posterior sample mean $\overline X_i'$ if $i\ne \alpha$, $B$ is the lower bound on $\overline X_i'$ if $i=\alpha$. For $i=\alpha$, the bound can be improved by selecting a midpoint
$B < y < \overline X_\beta$ and computing the bound as the sum of two parts:
\begin{itemize}
\item $\overline X_\beta-y$ multiplied by the probability that
  $\overline X_i' \le \overline X_\beta$;
\item $\overline X_\beta-B$ multiplied by the probability that $\overline X_i'\le
  y$.
\end{itemize}.
\begin{equation}
\Lambda_{i|i=\alpha} \le \hat \Lambda_{i|i=\alpha} =
 (\overline X_\beta-y)\exp\left(-\gamma n_i(\overline X_i-\overline X_\beta)^2\right)
 +(\overline X_\beta-B)\exp\left(-\gamma n_i(\overline X_i-y)^2\right)
\label{eqn:bound-twopart-hoeffding-alpha}
\end{equation}
If $y$ that minimizes $\hat\Lambda_{i|i=\alpha}$ exists, the minimum is achieved when
 $\frac {d\hat\Lambda_{i|i=\alpha}} {dy}=0$, that
is, when $y$ is the root of the following equation:
\begin{equation}
2\gamma n_i(\overline X_\beta-B)(\overline X_i-y)=\exp\left(-\gamma n_i\left(\frac {\overline X_i-\overline X_\beta}
    {\overline X_i-y}\right)^2\right)
\end{equation}
The derivation for the case $i\ne \alpha$ is obtained by substitution $1-A, 1-\overline X_i, 1-\overline X_\alpha, 1-y$ instead of
$B, \overline X_i, \overline X_\beta, y$ into (\ref{eqn:bound-twopart-hoeffding-alpha}):
\begin{equation}
\Lambda_{i|i\ne\alpha} \le \hat\Lambda_{i|i\ne\alpha} = (y-\overline X_\alpha)\exp\left(-\gamma n_i(\overline X_\alpha-\overline
  X_i)^2\right)+(A-\overline X_\alpha)\exp\left(-\gamma n_i(y-\overline X_i)^2\right)
\label{eqn:bound-twopart-hoeffding-rest}
\end{equation}
A closed-form solution for $y$ cannot be obtained, but given
$A, B, \overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
computed.

### 2011-12-07 paper/voisam/onbetterbound.tex
initial commit

Consider the case of VOI of measurement the current best item (sample mean
$\alpha$). The probability that the true mean of the item $\mu_\alpha$ is at
most $x$ is bounded by the Hoeffding's inequality as:
\[ \Pr(\mu_\alpha \le x) \le \exp\left(-2n(\alpha-x)^2\right) \]
An upper bound on the VOI is thus the product of the probability that
the true mean of the current best item is less than the sample mean
$\beta$ of the current second-best item times the maximum reward,
$\beta$:
\[ V \le \beta \exp\left(-2n(\alpha-\beta)^2\right) \]
The bound can be supposedly be improved by selecting a midpoint
$0 < \gamma < \beta$ and computing the bound as the sum of two parts:
\begin{itemize}
\item $\beta-\gamma$ multiplied by the probability that
  $\mu_\alpha \le \beta$;
\item $\beta$ multiplied by the probability that $\mu_\alpha\le
  \gamma$.
\end{itemize}.
\[ V^* \le (\beta-\gamma)\exp\left(-2n(\alpha-\beta)^2\right)+\beta\exp\left(-2n(\alpha-\gamma)^2\right) \]

@@

The derivation for the other case
(sampling an item that can be better than the current best) is
obtained by substitution $1-\alpha, 1-\beta, 1-\gamma$ instead of
$\alpha, \beta, \gamma$. The anticipated influence of the improved
VOI estimate would be that the selected item will be a less discovered
one and further from the current best or second-best.

>>>>

Consider the case of VOI of measurement the current best item (sample mean
$\alpha$). The probability that the true mean of the item $\mu_\alpha$ is at
most $x$ is bounded by the Hoeffding's inequality as:
\[ \Pr(\mu_\alpha \le x) \le \exp\left(-2n(\alpha-x)^2\right) \]
An upper bound on the VOI is thus the product of the probability that
the true mean of the current best item is less than the sample mean
$\beta$ of the current second-best item times the maximum reward,
$\beta$:
\[ V_{\overline x=\alpha} \le \beta \exp\left(-2n(\alpha-\beta)^2\right) \]
The bound can be supposedly be improved by selecting a midpoint
$0 < \gamma < \beta$ and computing the bound as the sum of two parts:
\begin{itemize}
\item $\beta-\gamma$ multiplied by the probability that
  $\mu_\alpha \le \beta$;
\item $\beta$ multiplied by the probability that $\mu_\alpha\le
  \gamma$.
\end{itemize}.
\[ V_{\overline x=\alpha} \le (\beta-\gamma)\exp\left(-2n(\alpha-\beta)^2\right)+\beta\exp\left(-2n(\alpha-\gamma)^2\right) \]

@@

The derivation for the other case
(sampling an item that can be better than the current best) is
obtained by substitution $1-\overline x, 1-\alpha, 1-\gamma$ instead of
$\alpha, \beta, \gamma$:

\[ V_{\overline x\ne\alpha} \le (\gamma-\alpha)\exp\left(-2n(\alpha-\overline
  x)^2\right)+(1-\alpha)\exp\left(-2n(\gamma-\overline x)^2\right) \]

The anticipated influence of the improved
VOI estimate would be that the selected item will be a less discovered
one and further from the current best or second-best.

### 2011-12-07 paper/voisam/paper.tex
initial commit

Theorem~4 in~\cite{MaurerPontil.benrstein} states that
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
Therefore
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
\delta.\]
$t=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
square equation
\[t^2-t\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge t)$,
gives
\[\Pr(\IE[X]-\overline X_n\ge t)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
Other derivations, giving slightly different results, are possible.

>>>>

Theorem~4 in~\cite{MaurerPontil.benrstein} states that
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
Therefore
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
\delta.\]
$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
square equation
\[a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge a)$,
gives
\[\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
Other derivations, giving slightly different results, are possible.
